---
title: "Pretty forest plots"
author: "Krista Karttunen"
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    theme: united
---

```{r setup, include=TRUE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)

library(tidyverse)
library(GGally)
library(tidybayes)
library(bayesplot)
library(posterior)
library(survival)
library(survminer)
library(ggfortify)
library(ggridges)
library(gridExtra)
library(tictoc)
library(knitr)
library(kableExtra)
library(skimr)
library(patchwork)
library(ggh4x) # For more flexible facet_grid2
library(purrr)
library(broom)
library(glue)
library(ggforce)
library(tableone)
library(UpSetR)
library(scales)
library(ggcorrplot)
library(tools)
library(htmltools)
library(webshot)
library(RColorBrewer)

save_figs <- TRUE
base <- "../training_models/training_with_snp1_group"
data_path <- "../../DATA/data_preprocessed_3_20250120"
fig_path <- sprintf("%s/figures", base)
table_path <- sprintf("%s/tables", base)
source("common.R")

```

**This Rmd is for producing clear and clean effect size plots for the thesis** 
**presentation and final thesis document. Here the model naming and titles** 
**can be modified accordingly.**

To be able to compare different modelling methods we want to see the effect sizes
next to each other.

The short names used in the model training codes should be modified for more
descriptive ones for the use for presenting the results:

- freq --> LogR
- freq_int --> LogRInt
- bayes --> BayesLogR
- spline --> BayesLogRSpline
- cox --> Cox
- tdc --> Tdc


## Loading cis list

Retrieving the cis list produced in the training_models.Rmd.

```{r Loading cis lists}
freq_cis_groups <- read_tsv(sprintf("%s/freq_groups_cis.tsv", table_path))
freq_int_cis_groups <- read_tsv(sprintf("%s/freq_int_groups_cis.tsv", table_path))
bayes_cis <- read_tsv(sprintf("%s/bayes_cis.tsv", table_path))
spline_cis <- read_tsv(sprintf("%s/spline_cis.tsv", table_path))
cox_cis <- read_tsv(sprintf("%s/cox_cis.tsv", table_path))
tdc_cis <- read_tsv(sprintf("%s/tdc_cis_train.tsv", table_path))
```

## Preparation for the plots (names, colors etc.)

```{r Preparation for plots}
groups <- c("all", "male", "female")

# Model names in the training_models.Rmd
# Change also the model_names, cis_value_list, model_titles, x_labs to the same ones
#mymodels <- c("freq","bayes_spline", "freq_int", "cox", "bayes", "tdc")

#Commend if just testing
mymodels <- c("LogR", "LogRInt", "BayesLogR", "BayesLogRSpline", "Cox", "Tdc")

model_names <- c(
  LogR = "LogR",
  LogRInt = "LogRInt",
  BayesLogR = "BayesLogR",
  BayesLogRSpline = "BayesLogRSpline",
  Cox = "Cox",
  Tdc = "Tdc"
)

cis_values_list <- list(
  LogR = freq_cis_groups,
  LogRInt = freq_int_cis_groups,
  BayesLogR = bayes_cis,
  BayesLogRSpline = spline_cis,
  Cox = cox_cis,
  Tdc = tdc_cis)

# model names for plot titles
# model_titles = c(
#   LogR = "Logistic regression",
#   LogRInt = "Logistic regression with interactions",
#   BayesLogR = "Bayesian logistic regression with interactions",
#   BayesLogRSpline = "Bayesian logR, splines for age, blood_volume, days_to_previous_fb",
#   Cox = "Cox proportional hazard model",
#   Tcd = "Cox with time-dependent variable for previous_Hb"
# )

model_colors <- c(
  LogR = "#66A61E",
  LogRInt = "#7570B3",
  BayesLogR = "#D95F02",
  BayesLogRSpline = "#1B9E77",
  Cox = "#E6AB02",
  Tdc = "#E7298A"
)

# Add this inside model_colors, if the original models names are used
  # freq = "#7FC97F",
  # freq_int = "#BEAED4",
  # bayes = "#FDC086",
  # spline = "#FFFF99",
  # cox = "#386CB0",
  # tdc = "#F0027F"

x_labs <- c(
  LogR = "Odds ratio",
  LogRInt = "Odds ratio", 
  BayesLogR = "Odds ratio", 
  BayesLogRSpline = "Odds ratio", 
  Cox = "Hazard ratio", 
  Tdc = "Hazard ratio")

```

## Combine the cis list into one table

```{r Creating groupped_tables}
# Add column "model" to all the data frames
freq_cis_groups$model <- "LogR"
freq_int_cis_groups$model <- "LogRInt"
bayes_cis$model <- "BayesLogR"
spline_cis$model <- "BayesLogRSpline"
cox_cis$model <- "Cox"
tdc_cis$model <- "Tdc"

# Combine the cis list and modify the Pretty labels for the interactions
combined_cis <-
  bind_rows(freq_cis_groups, spline_cis, freq_int_cis_groups, cox_cis,
            bayes_cis, tdc_cis) %>%
  mutate(Pretty = case_when(
    Pretty == "smokingTRUE:femaleTRUE" ~ "Smoking:Female",
    Pretty == "snp_1_169549811:grouppost_menopausal_female" ~ "SNP 1:169549811:Postmenopausal Female",
    Pretty == "snp_1_169549811:grouppre_menopausal_female" ~ "SNP 1:169549811:Premenopausal Female",
    Pretty == "snp_22_37066896:grouppost_menopausal_female" ~ "SNP 22:37066896:Postmenopausal female",
    Pretty == "snp_22_37066896:grouppre_menopausal_female" ~ "SNP 22:37066896:Premenopausal female",
    TRUE ~ Pretty))

# Create a groupped_tables were the cis values are divided by group
groupped_tables <- combined_cis %>%
  group_split(group) %>%
  set_names(c("all", "female", "male"))

#groupped_tables
```


## Separate forest plots for all demogrpahic groups

Creating forest plots of the effect sized separately for the groups.

```{r Effect sizes for all models in one - all, fig.dim=c(8,12)}

g <- groupped_tables$all %>% 
  mutate(Pretty = fct_rev(Pretty)) %>%
  mutate(Model = factor(model, levels = mymodels)) %>%
  mutate(hollow_model = factor(ifelse(low <= 1 & high >= 1, NA_character_, as.character(Model)), 
                               levels = mymodels)) %>%
  mutate(across(c(Model, hollow_model), fct_rev)) %>% # this turns the values the same way as the legend
  ggplot(aes(x = estimate, xmin = low, xmax = high, y = term, # y = Pretty for pretty labs
             color = Model, 
             fill = hollow_model)) + 
  geom_vline(aes(xintercept=1), color="gray", size=1) +
  stat_stripes(aes(stripe = term), alpha = 0.5, fill = "lightgray") +
  ggstance::geom_linerangeh(position = position_dodge(width = 1), size = 0.2) +
  geom_point(shape = 21, size = 1.5, position = position_dodge(width = 1)) +
  labs(title = "Effect sizes for group - all", x="Odds ratio/Hazard ratio", y="Standardized variables") + 
  scale_colour_manual(values = model_colors, labels = model_names, breaks = mymodels) +
  scale_fill_manual(guide = "none", values = model_colors, na.value = NA) +
  guides(color = guide_legend(override.aes = list(shape = 16))) +
  scale_x_log10() +
  scale_y_discrete() 

if (save_figs) {
  filename <- sprintf("%s/combined_effect_sizes_all.png", fig_path)
  ggsave(filename=filename,  plot=g, width = 210,  height = 297, units="mm", dpi=350, scale=1.0, device="png")
}
g
```

```{r Effect sizes for all models in one - male, fig.dim=c(8,12)}

g <- groupped_tables$male %>% 
  mutate(Pretty = fct_rev(Pretty)) %>%
  mutate(Model = factor(model, levels = mymodels)) %>%
  mutate(hollow_model = factor(ifelse(low <= 1 & high >= 1, NA_character_, as.character(Model)), 
                               levels = mymodels)) %>%
  mutate(across(c(Model, hollow_model), fct_rev)) %>% # this turns the values the same way as the legend
  ggplot(aes(x = estimate, xmin = low, xmax = high, y = term, # y = Pretty for pretty labs
             color = Model, 
             fill = hollow_model)) + 
  geom_vline(aes(xintercept=1), color="gray", size=1) +
  stat_stripes(aes(stripe = Pretty), alpha = 0.5, fill = "lightgray") +
  ggstance::geom_linerangeh(position = position_dodge(width = 1), size = 0.2) +
  geom_point(shape = 21, size = 1.5, position = position_dodge(width = 1)) +
  labs(title = "Effect sizes for group - male", x="Odds ratio/Hazard ratio", y="Standardized variables") + 
  scale_colour_manual(values = model_colors, labels = model_names, breaks = mymodels) +
  scale_fill_manual(guide = "none", values = model_colors, na.value = NA) +
  guides(color = guide_legend(override.aes = list(shape = 16))) +
  scale_x_log10() +
  scale_y_discrete() 

if (save_figs) {
  filename <- sprintf("%s/combined_effect_sizes_male.png", fig_path)
  ggsave(filename=filename,  plot=g, width = 180,  height = 350, units="mm", dpi=300, scale=1.0, device="png")
}
g

```

```{r Effect sizes for all models in one - female, fig.dim=c(8,12)}

g <- groupped_tables$female %>% 
  mutate(Pretty = fct_rev(Pretty)) %>%
  mutate(Model = factor(model, levels = mymodels)) %>%
  mutate(hollow_model = factor(ifelse(low <= 1 & high >= 1, NA_character_, as.character(Model)), 
                               levels = mymodels)) %>%
  mutate(across(c(Model, hollow_model), fct_rev)) %>% # this turns the values the same way as the legend
  ggplot(aes(x = estimate, xmin = low, xmax = high, y = term, # y = Pretty for pretty labs
             color = Model, 
             fill = hollow_model)) + 
  geom_vline(aes(xintercept=1), color="gray", size=1) +
  stat_stripes(aes(stripe = Pretty), alpha = 0.5, fill = "lightgray") +
  ggstance::geom_linerangeh(position = position_dodge(width = 1), size = 0.2) +
  geom_point(shape = 21, size = 1.5, position = position_dodge(width = 1)) +
  labs(title = "Effect sizes for group - female", x="Odds ratio/Hazard ratio", y="Standardized variables") + 
  scale_colour_manual(values = model_colors, labels = model_names, breaks = mymodels) +
  scale_fill_manual(guide = "none", values = model_colors, na.value = NA) +
  guides(color = guide_legend(override.aes = list(shape = 16))) +
  scale_x_log10() +
  scale_y_discrete() 

if (save_figs) {
  filename <- sprintf("%s/combined_effect_sizes_female.png", fig_path)
  ggsave(filename=filename,  plot=g, width = 180,  height = 350, units="mm", dpi=300, scale=1.0, device="png")
}
g
```


## Separate plots for non genomic and genomic variables all groups in one

I want to have two plots that include all groups (all, female, male):

1) Non genomic variables
2) Genomic variables (separated into two)

```{r Dividing variables into lists}
non_genomic_variables <- c("age", "blood_volume", "days_to_previous_fb",
                          "femaleTRUE", "previous_Hb", "recent_deferrals",
                          "recent_donations", "smokingTRUE", "smokingTRUE:femaleTRUE",
                          "warm_seasonTRUE", "year")

# There are so many genomic variables, it is better to have two separate plots for them
genomic_variables_1 <- c("snp_1_113834946", "snp_1_169549811",
                         "snp_1_169549811:grouppre_menopausal_female",
                         "snp_1_169549811:grouppost_menopausal_female",
                          "snp_4_25970243", "snp_6_25857692", "snp_6_32617727", "snp_6_32658525", 
                          "snp_6_68206710", "snp_7_75844637", "snp_8_10785723", "snp_8_23520397", 
                          "snp_8_75503352")

genomic_variables_2 <- c("snp_9_133271182", "snp_10_63174788", "snp_11_5226799", "snp_12_6030341",
                         "snp_14_33938877", "snp_15_45099877", "snp_17_58358769", "snp_20_35186730",
                         "snp_22_29795932", "snp_22_37066896", "snp_22_37066896:grouppost_menopausal_female",
                         "snp_22_37066896:grouppre_menopausal_female")
```


```{r Combined effect sizeds: non-genomic variables}

g <- combined_cis %>%
  filter(term %in% non_genomic_variables) %>%
  mutate(Pretty = fct_relevel(Pretty, function(levels) str_sort(levels, numeric=TRUE))) %>%
  filter(group %in% c("all", "male", "female")) %>%
  mutate(group = factor(group, levels=c("all", "male", "female"))) %>%
  mutate(model = factor(model, levels = mymodels)) %>%  # Set factor levels for model
  mutate(across(c(model, Pretty), fct_rev)) %>%
  mutate(hollow_model=factor(if_else(low<=1 & high>=1, NA_character_, as.character(model)),
                             levels=levels(model))) %>%
  ggplot(aes(x=estimate, xmin=low, xmax=high, y=Pretty, # y=Pretty for pretty labs, y=term for variable names, used in training
             color=model, fill=hollow_model)) + 
  geom_vline(aes(xintercept=1), color="gray", 
             #size=1
             linewidth=1
             ) +
  geom_stripes(freq_int_cis_groups %>% filter(term %in% non_genomic_variables), Pretty) +
  ggstance::geom_linerangeh(position=position_dodge(width=1), 
                            size=0.2
                            #linewidth=0.2
                            ) + 
  geom_point(#aes(), 
             #fill=NA, 
             shape=21,
             size=1.5,
             position=position_dodge(width=1)) +
  labs(#title = "Combined effect sizes: non-genomic", 
       x="Odds ratio/Hazards ratio", y="Standardized variables") + 
  scale_colour_manual(values = model_colors, labels = model_names, breaks = mymodels) +
  scale_fill_manual(guide = "none", values = model_colors, na.value = NA) +
  facet_wrap(~group) +
  guides(color = guide_legend(override.aes = list(shape = 16))) +
  #coord_cartesian(xlim = c(0.0001, 12)) +
  scale_x_log10() +
  scale_y_discrete() + # This is important. Solves the problem with position_dodge2 
                       # and the order of rect and pointrange geoms!!!
                       # Otherwise following error results: Error: Discrete value supplied to continuous scale
  theme(legend.position="bottom",
        legend.direction = "horizontal",
        plot.title = element_text(face="bold"),
        plot.subtitle = element_text(face="bold"))

if (save_figs) {
  # save as pdf
  # filename <- sprintf("%s/combined_effect_sizes_non_genomic.pdf", fig_path)
  # ggsave(filename=filename, plot=g, width = 300,  height=300, units="mm", dpi=300, scale=1.0, device="pdf")
  
  # save as png
  filename <- sprintf("%s/combined_effect_sizes_non_genomic.png", fig_path)
  ggsave(filename=filename, plot=g, width = 210,  height=250, units="mm", dpi=350, device="png")
}

g
```

```{r Combined effect sizes: genomic variables 1}
g <- combined_cis %>%
  filter(term %in% genomic_variables_1) %>% 
  mutate(Pretty = fct_relevel(Pretty, function(levels) str_sort(levels, numeric=TRUE))) %>%
  filter(group %in% c("all", "male", "female")) %>%
  mutate(group = factor(group, levels=c("all", "male", "female"))) %>%
  mutate(model = factor(model, levels = mymodels)) %>% 
  mutate(across(c(model, Pretty), fct_rev)) %>%
  mutate(hollow_model=factor(if_else(low<=1 & high>=1, NA_character_, as.character(model)),
                             levels=levels(model))) %>%
  ggplot(aes(x=estimate, xmin=low, xmax=high, y=Pretty, # y=Pretty for pretty labs, y=term for variable names, used in training
             color=model, fill=hollow_model)) + 
  geom_vline(aes(xintercept=1), color="gray", 
             #size=1
             linewidth=1
             ) +
  geom_stripes(freq_int_cis_groups %>% filter(term %in% genomic_variables_1), Pretty) +
  ggstance::geom_linerangeh(position=position_dodge(width=1), 
                            size=0.2
                            #linewidth=0.2
                            ) + 
  geom_point(#aes(), 
             #fill=NA, 
             shape=21,
             size=1.5,
             position=position_dodge(width=1)) +
  labs(#title = "Combined effect sizes: genomic", 
       x="Odds ratio/Hazards ratio", y="Standardized variables") + 
  scale_colour_manual(values = model_colors, labels = model_names, breaks = mymodels) +
  scale_fill_manual(guide = "none", values = model_colors, na.value = NA) +
  facet_wrap(~group) +
  guides(color = guide_legend(override.aes = list(shape = 16))) +
  #coord_cartesian(xlim = c(0.0001, 12)) +
  scale_x_log10() +
  scale_y_discrete() + # This is important. Solves the problem with position_dodge2 
                       # and the order of rect and pointrange geoms!!!
                       # Otherwise following error results: Error: Discrete value supplied to continuous scale
  theme(legend.position="bottom",
        legend.direction = "horizontal",
        plot.title = element_text(face="bold"),
        plot.subtitle = element_text(face="bold"))

if (save_figs) {
  #save as pdf
  # filename <- sprintf("%s/combined_effect_sizes_genomic.pdf", fig_path)
  # ggsave(filename=filename, plot=g, width = 300,  height=350, units="mm", dpi=300, scale=1.0, device="pdf")
  
  #save as png
  filename <- sprintf("%s/combined_effect_sizes_genomic_1.png", fig_path)
  ggsave(filename=filename, plot=g, width = 210,  height=250, units="mm", dpi=350, device="png")
}

g
```

```{r Combined effect sizes: genomic variables 2}
g <- combined_cis %>%
  filter(term %in% genomic_variables_2) %>%
  mutate(Pretty = fct_relevel(Pretty, function(levels) str_sort(levels, numeric=TRUE))) %>%
  filter(group %in% c("all", "male", "female")) %>%
  mutate(group = factor(group, levels=c("all", "male", "female"))) %>%
  mutate(model = factor(model, levels = mymodels)) %>% 
  mutate(across(c(model, Pretty), fct_rev)) %>%
  mutate(hollow_model=factor(if_else(low<=1 & high>=1, NA_character_, as.character(model)),
                             levels=levels(model))) %>%
  ggplot(aes(x=estimate, xmin=low, xmax=high, y=Pretty, # y=Pretty for pretty labs, y=term for variable names, used in training
             color=model, fill=hollow_model)) + 
  geom_vline(aes(xintercept=1), color="gray", 
             #size=1
             linewidth=1
             ) +
  geom_stripes(freq_int_cis_groups %>% filter(term %in% genomic_variables_2), Pretty) +
  ggstance::geom_linerangeh(position=position_dodge(width=1), 
                            size=0.2
                            #linewidth=0.2
                            ) + 
  geom_point(#aes(), 
             #fill=NA, 
             shape=21,
             size=1.5,
             position=position_dodge(width=1)) +
  labs(#title = "Combined effect sizes: genomic", 
       x="Odds ratio/Hazards ratio", y="Standardized variables") + 
  scale_colour_manual(values = model_colors, labels = model_names, breaks = mymodels) +
  scale_fill_manual(guide = "none", values = model_colors, na.value = NA) +
  facet_wrap(~group) +
  guides(color = guide_legend(override.aes = list(shape = 16))) +
  #coord_cartesian(xlim = c(0.0001, 12)) +
  scale_x_log10() +
  scale_y_discrete() + # This is important. Solves the problem with position_dodge2 
                       # and the order of rect and pointrange geoms!!!
                       # Otherwise following error results: Error: Discrete value supplied to continuous scale
  theme(legend.position="bottom",
        legend.direction = "horizontal",
        plot.title = element_text(face="bold"),
        plot.subtitle = element_text(face="bold"))

if (save_figs) {
  # save as pdf
  # filename <- sprintf("%s/combined_effect_sizes_genomic.pdf", fig_path)
  # ggsave(filename=filename, plot=g, width = 300,  height=350, units="mm", dpi=300, scale=1.0, device="pdf")
  
  #save as png
  filename <- sprintf("%s/combined_effect_sizes_genomic_2.png", fig_path)
  ggsave(filename=filename, plot=g, width = 210,  height=250, units="mm", dpi=350, device="png")
}

g
```

### Plotting significant genomic variables

Plotting only the variables that are significant in any of the groups or models.

Here the significant genomic variables are hand listed, but a method to pick them
from the combined_cis or groupped_tables. 

```{r Creating list of signifcant genomic variables}
significant_genomic <- c("snp_1_113834946", "snp_1_169549811",
                         "snp_1_169549811:grouppre_menopausal_female",
                         "snp_1_169549811:grouppost_menopausal_female",
                         "snp_6_25857692", "snp_10_63174788", "snp_11_5226799",
                         "snp_12_6030341", "snp_17_58358769", "snp_20_35186730",
                          "snp_22_37066896", 
                         "snp_22_37066896:grouppost_menopausal_female",
                         "snp_22_37066896:grouppre_menopausal_female")
```

```{r}
g <- combined_cis %>%
  filter(term %in% significant_genomic) %>%
  mutate(Pretty = fct_relevel(Pretty, function(levels) str_sort(levels, numeric=TRUE))) %>%
  filter(group %in% c("all", "male", "female")) %>%
  mutate(group = factor(group, levels=c("all", "male", "female"))) %>%
  mutate(model = factor(model, levels = mymodels)) %>% 
  mutate(across(c(model, Pretty), fct_rev)) %>%
  mutate(hollow_model=factor(if_else(low<=1 & high>=1, NA_character_, as.character(model)),
                             levels=levels(model))) %>%
  ggplot(aes(x=estimate, xmin=low, xmax=high, y=Pretty, # y=Pretty for pretty labs
             color=model, fill=hollow_model)) + 
  geom_vline(aes(xintercept=1), color="gray", 
             #size=1
             linewidth=1
             ) +
  geom_stripes(freq_int_cis_groups %>% filter(term %in% significant_genomic), Pretty) +
  ggstance::geom_linerangeh(position=position_dodge(width=1), 
                            size=0.2
                            #linewidth=0.2
                            ) + 
  geom_point(#aes(), 
             #fill=NA, 
             shape=21,
             size=1.5,
             position=position_dodge(width=1)) +
  labs(#title = "Combined effect sizes: genomic", 
       x="Odds ratio/Hazard ratio", y="Standardized variables") + 
  scale_colour_manual(values = model_colors, labels = model_names, breaks = mymodels) +
  scale_fill_manual(guide = "none", values = model_colors, na.value = NA) +
  facet_wrap(~group) +
  guides(color = guide_legend(override.aes = list(shape = 16))) +
  #coord_cartesian(xlim = c(0.0001, 12)) +
  scale_x_log10() +
  scale_y_discrete() + # This is important. Solves the problem with position_dodge2 
                       # and the order of rect and pointrange geoms!!!
                       # Otherwise following error results: Error: Discrete value supplied to continuous scale
  theme(legend.position="bottom",
        legend.direction = "horizontal",
        plot.title = element_text(face="bold"),
        plot.subtitle = element_text(face="bold"))

if (save_figs) {
  # save as pdf
  # filename <- sprintf("%s/combined_effect_sizes_genomic.pdf", fig_path)
  # ggsave(filename=filename, plot=g, width = 300,  height=350, units="mm", dpi=300, scale=1.0, device="pdf")
  
  # save as png
  filename <- sprintf("%s/combined_effect_sizes_significant_genomic.png", fig_path)
  ggsave(filename=filename, plot=g, width = 210,  height=250, units="mm", dpi=350, device="png")
}

g
```

